```{r}
setwd('/home/lien/data/skrypty/TFIIB/')
library(tidyverse)
source('functions.R')

```
# provide the path to GMatrix.RDS files
```{r}

data_dir='/home/lien/data/TFIIB/GEO_upload/HiC/'

sample_info <- read_tsv('/home/lien/data/skrypty/TFIIB/samples_GEO_HiC.tsv',col_types='ffffff')
group_sets=list(
  'other'=c('All_PCG'))
down_thresh=3000
up_thresh=50000
samples=c(1:12)
Groups <- create_Groups(group_sets,down_thresh,up_thresh)
```


#* if the meta-gene matrixes for 'All_PCG' were already calculated and saved, for example by running reproduce_Fig3E.R, the following part can be skipped (it takes a lot of time)


Create a unique output directory
```{r}
barcode=paste0(paste0(sample(c(letters,LETTERS),5,replace=T),collapse=""),as.integer(as.POSIXct(Sys.time())))
output_path = paste0(barcode,'/')
system(paste0('mkdir ',output_path),intern=F)
```


Calculate meta-gene matrizes for specified groups and samples. Immediately save the results as symmetric matrices.
Both raw and distance-normalized matrixes are generated
```{r}
system.time({metamatrix_data <- lapply(1:length(Groups),function(G){
  lapply(samples,function(s){
    genes=Groups[[G]]
    result <- calculate_metamatrix(genes,s,reach=down_thresh/2)
    result$sample=paste0('Sample',s)
    result$name <- unite(sample_info,c(2:4),col='name',sep='_' )[which(sample_info$sample==result$sample),] %>% pull(name)
    result$group=names(Groups)[G]
    title=paste0(result$sample,'_',result$group,'_',down_thresh/1000,' kb to ',up_thresh/1000,' kb.matrix')
    write.table(result$mat_raw, file=paste0(output_path,'Meta_raw_',title), sep='\t', row.names = F, col.names =  F)
    write.table(result$mat_norm, file=paste0(output_path,'Meta_norm_',title), sep='\t', row.names = F, col.names =  F)
    return(result)
  })
}) %>% reduce(c)})
```


# plot the first gene group from the first sample
```{r}
x=metamatrix_data[[1]]
title=paste0(x$sample,'   ',x$name,'\n',x$group,'; ',down_thresh/1000,' kb to ',up_thresh/1000,' kb\n',x$nr_of_genes,' genes')
f_draw_matrix(raw_array=x$mat_raw, norm_array=x$mat_norm,title=title,threshold=0.995)

```

# create a pdf report with all heatmaps
```{r}
pdf_file=paste0(output_path,paste0(names(Groups),collapse='_'),'.pdf')
pdf(pdf_file)
lapply(metamatrix_data, function(x){

  title=paste0(x$sample,'   ',x$name,'\n',x$group,'; ',down_thresh/1000,' kb to ',up_thresh/1000,' kb\n',x$nr_of_genes,' genes')
  f_draw_matrix(raw_array=x$mat_raw, norm_array=x$mat_norm,title=title,threshold=0.995)

})
dev.off()
```

############ heatmaps from the Fig. 3 E - combine biological repeats for each condition 
############ and differential maps TFIIB1R58C-TFIIB1-WT
###### draw heatmaps from saved matrix files (to get the common scale)
# set path to precomputed meta_gene matrixes for 'All_PCG' all protein coding genes longer than 3kb
```{r}
output_path = '/home/lien/data/skrypty/TFIIB/Foiki1648740534/'
Samples=list(
  WT=list(Water=c(1,9),S1=c(2,10),Recovery=c(3,11),S2=4),
  R58C=list(Water=5,S1=c(6,12),Recovery=7, S2=8))
matrices <- lapply(names(Groups),function(group){
  # group=names(Groups)[3]
   lapply(Samples,function(genotype){
    lapply(genotype,function(condition){
      lapply(condition,function(sample){
        norm_matrix <- read_tsv(paste0(output_path,'Meta_norm_Sample',sample,'_',group,'_',down_thresh/1000,' kb to ',up_thresh/1000,' kb.matrix'), col_names=F,col_types = cols(.default = "d")) %>%
          as.matrix()
         scaled_norm_matrix <- (norm_matrix-mean(norm_matrix,na.rm=T))/sd(norm_matrix,na.rm=T)
      }) %>% Reduce('+',.)/length(condition)
    })
  })
})
names(matrices)<-names(Groups)
  
 
  
conditions=c('Water','S1','Recovery','S2')
deltas <- lapply(names(Groups),function(group){lapply(conditions,function(condition){
 d_array = matrices[[group]][['R58C']][[condition]]-matrices[[group]][['WT']][[condition]]
 title=paste0(group,'\n',condition,'\nR58C-WT\nDifferential matrix')
 return(list(delta_array=d_array, group=group, condition=condition))
 })})%>% reduce(c)

```

# set the common scale for heatmaps
 
```{r}
threshold=0.999
max_norm = quantile(abs((unlist(matrices))),probs=threshold,na.rm=T)
max_delta = quantile(abs((unlist(lapply(deltas,function(x){x$delta_array})))),probs=threshold,na.rm=T)
```

# plot the examplary set of heatmaps (All protein coding genes, Stress1, TFIIB1-WT and TFIIB1-R58C):
```{r}
group='All_PCG'
lapply(names(Samples),function(genotype){
condition='S1'
    norm_map <- matrices[[group]][[genotype]][[condition]]
    title=paste0(group,' \nThresh: ',threshold,'\n',genotype,' ',condition)
    
    f_draw_matrix(norm_array=norm_map, title=title, do_scaling=F,max_scale_norm=max_norm)
})
d=deltas[[2]]
  title=paste0(d$group,'\n',d$condition,'\nR58C-WT')
  f_draw_matrix(delta_array=d$delta_array, title=title, do_scaling=F, max_scale_delta=max_delta)



```

# make a report with all heatmaps
```{r }
  pdf_file=paste0(output_path,paste0(names(Groups),collapse='_'),'comb_replicates.pdf')
  pdf(pdf_file)
lapply(names(Groups),function(group){ 
lapply(names(Samples),function(genotype){
  lapply(conditions,function(condition){
    norm_map <- matrices[[group]][[genotype]][[condition]]
    title=paste0(group,' \nThresh: ',threshold,'\n',genotype,' ',condition)
    
    f_draw_matrix(norm_array=norm_map, title=title, do_scaling=F,max_scale_norm=max_norm)
  })
})
})
lapply(deltas, function(d){
  title=paste0(d$group,'\n',d$condition,'\nR58C-WT')
  
  f_draw_matrix(delta_array=d$delta_array, title=title, do_scaling=F, max_scale_delta=max_delta)
 })
  dev.off()
```


############# heatmaps for the Supplemental Fig. 3 C - individual biological repeats, common scale
############ and differential maps TFIIB1R58C-TFIIB1-WT
###### draw heatmaps from saved matrix files (to get the common scale)
# set path to precomputed meta_gene matrixes
```{r}
# output_path = ''
matrices <-  lapply(names(Groups), function(group){
  lapply(samples,function(sample){
    norm_matrix <- read_tsv(paste0(output_path,'Meta_norm_Sample',sample,'_',group,'_',down_thresh/1000,' kb to ',up_thresh/1000,' kb.matrix'), col_names=F, col_types = cols(.default = "d")) %>%
      as.matrix()
    scaled_norm_matrix <- (norm_matrix-mean(norm_matrix,na.rm=T))/sd(norm_matrix,na.rm=T)
    
  })
})
names(matrices)=names(Groups) 
```


# set the common scale for heatmaps
```{r}
threshold=0.995
max_norm = quantile(abs((unlist(matrices))),probs=threshold,na.rm=T)
```


# plot the first heatmap
```{r}
group=names(Groups)[1]
sample=samples[1]
name <- unite(sample_info,c(2:4),col='name',sep='_' )[which(sample_info$sample==paste0('Sample',sample)),] %>% pull(name)
norm_map <- matrices[[group]][[sample]]
    title=paste0(group,' \nThresh: ',threshold,'\nSample ',sample,'   ',name)
    f_draw_matrix(norm_array=norm_map, title=title, do_scaling=F, max_scale_norm=max_norm)
```


# make a report with all heatmaps
```{r}
pdf_file=paste0(output_path,paste0(names(Groups),collapse='_'),'ind_replicates.pdf')
pdf(pdf_file)
lapply(names(Groups),function(group){ 
  lapply(samples,function(sample){
    norm_map <- matrices[[group]][[sample]]
    name <- unite(sample_info,c(2:4),col='name',sep='_' )[which(sample_info$sample==paste0('Sample',sample)),] %>% pull(name)
    title=paste0(group,' \nThresh: ',threshold,'\nSample ',sample,'   ',name)
    f_draw_matrix(norm_array=norm_map, title=title, do_scaling=F, max_scale_norm=max_norm)
  })
})

dev.off()
```

