---
output:
  html_document: default
  pdf_document: default
---
```{r}
setwd('/home/lien/data/skrypty/TFIIB/')
library("tidyverse")
source('functions.R')
```


#####  provide the path to GMatrix.RDS files
```{r}
data_dir='/home/lien/data/TFIIB/GEO_upload/HiC/'
```


#####  load the metadata for Hi-C samples
```{r}
sample_info <- read_tsv('/home/lien/data/skrypty/TFIIB/samples_GEO_HiC.tsv',col_types='ffffff')
sample_info
```


#####  analyse all samples
```{r}
samples=c(1:12)

```


#####  prepare the lists of genes: from 3 to 50kb, that lost or gained trainability in TFIIB1-R58C  (see functions.R and reproduce_Fig1and2.Rmd for details).
```{r}
group_sets=list('other'=c('Trainability_Loss','Trainability_Gain'))
down_thresh=3000
up_thresh=50000
Groups <- create_Groups(group_sets,down_thresh,up_thresh,complement_rest=F,exclusive=F)

```


#####  Create a unique directory to save matrix files and reports
```{r}
barcode=paste0(paste0(sample(c(letters,LETTERS),5,replace=T),collapse=""),as.integer(as.POSIXct(Sys.time())))
output_path = paste0(barcode,'/')
system(paste0('mkdir ',output_path),intern=F)
```


#####  Calculate meta-gene matrizes for specified groups and samples. Immediately save the results as symmetric matrices.
#####  Both raw and distance-normalized matrixes are generated.
```{r}
system.time({
metamatrix_data <- lapply(1:length(Groups),function(G){
  lapply(samples,function(s){
    genes=Groups[[G]]
    result <- calculate_metamatrix(genes,s)
    result$sample=paste0('Sample',s)
    result$name <- unite(sample_info,c(2:4),col='name',sep='_' )[which(sample_info$sample==result$sample),] %>% pull(name)
    result$group=names(Groups)[G]
    title=paste0(result$sample,'_',result$group,'_',down_thresh/1000,' kb to ',up_thresh/1000,' kb.matrix')
    write.table(result$mat_raw, file=paste0(output_path,'Meta_raw_',title), sep='\t', row.names = F, col.names =  F)
    write.table(result$mat_norm, file=paste0(output_path,'Meta_norm_',title), sep='\t', row.names = F, col.names =  F)
    return(result)
  })
}) %>% reduce(c)
})

```

#####  Plot the first gene group from the first sample
```{r}
x=metamatrix_data[[1]]
title=paste0(x$sample,'   ',x$name,'\n',x$group,'; ',down_thresh/1000,' kb to ',up_thresh/1000,' kb\n',x$nr_of_genes,' genes')
f_draw_matrix(raw_array=x$mat_raw, norm_array=x$mat_norm,title=title,threshold=0.99)

```

#### Supplemental Fig. 5
#####  set the common scale for heatmaps
```{r}
threshold=0.995
max_norm = quantile(abs(scale(unlist(lapply(metamatrix_data,function(x){x$'mat_norm'})))),probs=threshold,na.rm=T)

```



##### create a pdf report with all heatmaps
```{r}
 pdf_file=paste0(output_path,paste0(unlist(group_sets),collapse='_'),'_commonScale.pdf')
 pdf(pdf_file)
lapply(metamatrix_data, function(x){
  title=paste0(x$sample,'   ',x$name,'\n',x$group,'; ',down_thresh/1000,' kb to ',up_thresh/1000,' kb\n',x$nr_of_genes,' genes')
     f_draw_matrix(raw_array=x$mat_raw, norm_array=x$mat_norm,title=title,threshold=threshold,max_scale_norm=max_norm)
})
dev.off()
```
####  Fig. 5, upper panels
#####  the heatmaps on Fig. 5 are from combined samples from water/drought conditions
#####  matrixes for each group of genes and for each sample can be loaded directly. Change the output_path to the dir containing matrices saved in the previous step, if necessary

```{r}
#output_path='uawFz1649059942/'
Samples=list(WT=list(water=c(1,3,9,11),drought=c(2,4,10)),R58C=list(water=c(5,7),drought=c(6,8,12)))
matrices <- lapply(names(Groups),function(group){
   lapply(Samples,function(genotype){
    lapply(genotype,function(condition){
      lapply(condition,function(sample){
        norm_matrix <- read_tsv(paste0(output_path,'Meta_norm_Sample',sample,'_',group,'_',down_thresh/1000,' kb to ',up_thresh/1000,' kb.matrix'), col_names=F, col_types=cols(.default = "d")) %>%
          as.matrix()
         scaled_norm_matrix <- (norm_matrix-mean(norm_matrix,na.rm=T))/sd(norm_matrix,na.rm=T)
      }) %>% Reduce('+',.)/length(condition)
    })
  })
})
names(matrices)<-names(Groups)
```

#####  calculate differential matrices (TFIIB1-WT - TFIIB1-R58C) in each condition
```{r}
conditions=c('drought','water')
deltas <- lapply(names(Groups),function(group){lapply(conditions,function(condition){
 d_array = matrices[[group]][['R58C']][[condition]]-matrices[[group]][['WT']][[condition]]
 title=paste0(group,'\n',condition,'\nR58C-WT\nDifferential matrix')
 return(list(delta_array=d_array, group=group, condition=condition))
 })})%>% reduce(c)
```


#####  set the common scale for heatmaps
```{r}
threshold=0.995
max_norm = quantile(abs((unlist(matrices))),probs=threshold,na.rm=T)
max_delta = quantile(abs((unlist(lapply(deltas,function(x){x$delta_array})))),probs=threshold,na.rm=T)

```
#####  plot the heatmaps
```{r}
lapply(names(Groups),function(group){ 
lapply(names(Samples),function(genotype){
  lapply(conditions,function(condition){
    norm_map <- matrices[[group]][[genotype]][[condition]]
    title=paste0(group,' \nThresh: ',threshold,'\n',genotype,' ',condition)
    
    f_draw_matrix(norm_array=norm_map, title=title, do_scaling=F,max_scale_norm=max_norm)
  })
})
})
lapply(deltas, function(d){
  title=paste0(d$group,'\n',d$condition,'\nR58C-WT')
  
  f_draw_matrix(delta_array=d$delta_array, title=title, do_scaling=F, max_scale_delta=max_delta)
})

```
#####  make a pdf report
```{r}
  pdf_file=paste0(output_path,paste0(names(Groups),collapse='_'),'_combSamples.pdf')
  pdf(pdf_file)
lapply(names(Groups),function(group){ 
lapply(names(Samples),function(genotype){
  lapply(conditions,function(condition){
    norm_map <- matrices[[group]][[genotype]][[condition]]
    title=paste0(group,' \nThresh: ',threshold,'\n',genotype,' ',condition)
    
    f_draw_matrix(norm_array=norm_map, title=title, do_scaling=F,max_scale_norm=max_norm)
  })
})
})
lapply(deltas, function(d){
  title=paste0(d$group,'\n',d$condition,'\nR58C-WT')
  
  f_draw_matrix(delta_array=d$delta_array, title=title, do_scaling=F, max_scale_delta=max_delta)
})
  dev.off()
```

####  Fig. 5, lower panels - mean asymmetry ratios in each sample
#####  Pull the data from *sectors.tsv files
```{r}

sector_data <- lapply(1:12,function(i){
  read_tsv(paste0('/home/lien/data/TFIIB/GEO_upload/HiC/Sample',i,'_sectors.tsv'), col_types='cdddc')
}) %>% reduce(bind_rows) 

```

#####  calculate mean asymmetry ratio for each group of genes, for each sample
```{r message=FALSE}
summary <- make_contact_summary(data=sector_data,Groups)
```

#####  create plots 
```{r}
plots <- report_contact_summary(summary,Groups)
```
#####  add a single plot with mean asymmetry ratios for different groups, can be averaged across multiple samples. Here, TFIIB1-WT non-treated samples are chosen (2 biological repeats)
```{r}
plots$boot_plot <- bootANDplot(data=sector_data,Groups,merge_samples=c(1,9))

lapply(plots,function(x){x})
```

#####  make a pdf report with all plots and table-summary
```{r}
pdf_name=paste0(output_path, 'AsymRatio_Fig5.pdf')
pdf(pdf_name,width = 4, height = 4)
print(plots)
lapply(names(Groups),function(x){
  table <- summary %>% select(group, sample,mean_AsRatio) %>%
    subset(group==x)%>%
    tableGrob(theme=ttheme_default(base_size=6))
   grid.newpage()
  grid.draw(table)
})

dev.off()
#####  plots can be saved as .svg
#####  lapply(1:length(plots),function(x){
#####    svg(paste0(output_path,x,'.svg'),width = 4, height = 4)
#####    print(plots[[x]])
#####    dev.off()
#####  })
```
#### Supplemental Fig.5B - single gene contact matrices
##### generate lists of genes
```{r}
group_sets=list('custom_gene_list'=c('AT1G03080','AT3G12580','AT5G20900','AT4G39980','AT2G39800',
                                     'AT1G09970','AT1G62660','AT4G16760',
                                     'AT4G18950','AT4G23050','AT5G10650','AT5G51070'))
#group_sets=list('other'=c('Trainability_Loss','Trainability_Gain'))
Groups <- create_Groups(group_sets)
```

##### define samples to combine data. Here, TFIIB1-WT and TFIIB1-R58C drought samples (see <sample_info>)  
```{r}
sample_sets=list('Wt'=as.character(c(2,4,10)),'R58C'=as.character(c(6,8,12)))

```

#####  extract a contact matrix for each gene, each sample
```{r}
data <- lapply(names(Groups),function(group){
  genes = Groups[[group]]
  d <- lapply(names(sample_sets),function(genotype){
    d <-lapply(sample_sets[[genotype]],function(sample){
      # browser()
      gene_matrix <- singleGene_map(genes,sample,format='norm')
    })
    names(d) <- sample_sets[[genotype]]
    return(d)
  })
  names(d) <- names(sample_sets)
  return(d)
})
names(data) <- names(Groups)

```

##### set the threshold for heatmaps common scale
```{r}
threshold=0.995
```


##### average tha matrixes for each gene across the sample sets 
```{r}
combined_repl_data <- lapply(names(Groups),function(group){
  d <-lapply(Groups[[group]],function(gene){
    d <-lapply(names(sample_sets), function(set){
      lapply(sample_sets[[set]], function(sample){
        norm_matrix <- data[[group]][[set]][[sample]][[gene]]
        scaled_norm_matrix <- (norm_matrix-mean(norm_matrix,na.rm=T))/sd(norm_matrix,na.rm=T)
      })%>% Reduce('+',.)/length(sample_sets[[set]])
    })
    names(d) <- names(sample_sets)
    d$max_norm = quantile(abs((unlist(d))),probs=threshold,na.rm=T)
    return(d)
  })
  names(d) <- Groups[[group]]
  return(d)
})
names(combined_repl_data) <- names(Groups)
```

##### plot the first gene:
```{r}
group=names(Groups)[1]
gene=Groups[[1]][1]
lapply(names(sample_sets), function(set){
  matrix <- combined_repl_data[[group]][[gene]][[set]]
  thresh=combined_repl_data[[group]][[gene]][['max_norm']]
  title=paste0(gene,' from group: ',group,'\n',set,' (',paste0(sample_sets[[set]],collapse=', '),')')
  f_draw_matrix(norm_array=matrix,title=title, max_scale_norm=thresh)
})

```

##### plot all heatmaps in a pdf report, 1 file for 1 group of genes defined in group_sets
```{r}
barcode=paste0(paste0(sample(c(letters,LETTERS),5,replace=T),collapse=""),as.integer(as.POSIXct(Sys.time())))
output_path = paste0(barcode,'_single_gene_matrices/')
system(paste0('mkdir ',output_path),intern=F)
lapply(1:length(Groups), function(g){
  group=names(Groups)[g]
  genes=Groups[[g]]
  pdf_file=paste0(output_path,group,'_G',g,'_single_metagenMaps.pdf')
  pdf(pdf_file)
  
  lapply(genes,function(gene){
    lapply(names(sample_sets), function(set){
      matrix <- combined_repl_data[[group]][[gene]][[set]]
      thresh=combined_repl_data[[group]][[gene]][['max_norm']]
      title=paste0(gene,' from group: ',group,'\n',set,' (',paste0(sample_sets[[set]],collapse=', '),')')
      f_draw_matrix(norm_array=matrix,title=title, max_scale_norm=thresh)
    })
  })
  dev.off() 
})
```

